### 2016-08-19 rebuttal.tex

\question The sentence ``SymPy expressions are immutable trees of Python objects.'' doesn't belong in the conclusion. This can be moved to the appropriate place when discussing SymPy's architecture.
\begin{solution}
This paper focuses on the architecture of SymPy, and we consider this to be a
core feature of SymPy's architecture, hence, it's inclusion in the conclusion.
We have also discussed the different points in this sentence in-depth in the architecture section.
\end{solution}

\question All ``submodules'' should be replaced by ``modules'' (examples:
lines 657, 662).
\label{rev1point37}
\begin{solution}
We are unclear how the reviewer is making distinctions between ``modules'',
``submodules'', and ``packages''. Our use coincides with the definitions used
in the official Python documentation (see
\url{https://docs.python.org/3/tutorial/modules.html#packages} and
\url{https://docs.python.org/3/glossary.html#term-package}). The distinction
between (sub)module and (sub)package in terms of Python libraries seems very
technical. To clarify things, we have modified the manuscript to use
``package'' only when referring to a library external to SymPy (such as
mpmath and Xy-pic).

To avoid confusion, and to be consistent with the definitions used by the
official Python documentation, we have modified the manuscript to use
``submodule'' to refer to any module that is a dotted submodule of
\texttt{sympy}, such as \texttt{sympy.physics.quantum}. This is to emphasize
that these modules all live within the \texttt{sympy} namespace, as ``module''
can suggest a separate library.
\end{solution}

\question The sentences in lines 657--661 should swap places, because ``areas of mathematics'' are discrete mathematics, concrete mathematics, etc., while simplifying expressions, performing common calculus operations, pretty printing expressions, etc.\ belongs to common operations (``other areas'' is also fine, albeit slightly wrong).
\begin{solution}
The ``areas of mathematics'' was intended to encompass all mathematical
operations. We have rewritten the first sentence as
\begin{quote}
SymPy supports a wide array of mathematical facilities.
\end{quote}
We have also changed ``other included areas'' to ``other supported facilities'' in the third sentence.

The order of the subsequent sentences was chosen to match the order they were
presented in the paper, so we have left them as-is.
\end{solution}

### 2016-08-20 rebuttal.tex

\section{Comments from Editor}

\begin{questions}
\question I found the \url{http://live.sympy.org/} site very convenient for trying things
out in SymPy, but it is only mentioned in the supplement, which might be
overlooked by readers. I suggest moving that section into the main paper, as
it provides a great way to play along with the examples while reading the
paper.
\begin{solution}
This sentence was added to the introduction:
\begin{quote}
All examples could be tested on the \href{http://live.sympy.org}{SymPy
Live} instance, that is an online Python shell, which uses the Google
App Engine to execute SymPy code.
\end{quote}
\end{solution}
\question The Basic Usage section omits what I think is an important point:
how does one distinguish between evaluating exp(1) in Python and exp(1) in
SymPy? In other words, how are symbolic constants specified?
\begin{solution}
  It's unclear what you are asking for. Both Python standard library and SymPy
  have exp callable, so it depends which symbols are imported in your global
  namespace: the \texttt{exp} function from the \texttt{math} module or the
  \texttt{exp} class from SymPy. In the later case a special symbolic constant
  will be produced.

We have added a sentence to the section 2.2, that shows symbolic
constants as an example of leaf nodes:
\begin{quote}
Symbols or symbolic constants, like $e$ or $\pi$, are other examples of
leaf nodes.
\begin{verbatim}
>>> exp(1)
E
>>> exp(1).args
()
>>> x.args
()
\end{verbatim}
\end{quote}
\end{solution}
\question The first
thing I tried after seeing Table 2 (simplication functions) did not work:

\begin{verbatim}
>>> trigsimp (exp( Matrix(2, 2, [0, -y, y, 0])) )
\end{verbatim}
fails to recognize cos and
sin. Am I expecting too much?
\begin{solution}
  This indeed does not work. The implementation in \verb|trigsimp| primarily
  thus far has focused on transforming trigonometric functions into other
  trigonometric functions. The ability to simplify complex exponentials is
  something that we would like to work.
  \href{https://github.com/sympy/sympy/issues/11459}{Issue 11459} in our
  public issue tracker tracks this feature.
\end{solution}
\question Section 3.6: are eigenvalues and singular
values included? If not, why not?
\begin{solution}
Yes, they are included.  The second example in that section addresses the computation of
eigenvalues.  Singular values can be computed with
\texttt{A.singular\_values()}. We have not included this example, as the
singular value expressions for our example matrix \texttt{A} are quite large,
but we have added ``singular values'' to the list of features in the second
paragraph of the section.
\end{solution}
\question Like one of the referees, I expected to
see MPFR mentioned in Section 4.1, at least to mention the pros and cons and
say why it isn't used.
\begin{solution}
Major advantage is that mpmath is a pure Python library.  It's slow c.f.
MPFR, but not too much to evaluate any reasonably complicated single
expression at interactive speed, which covers most use cases in SymPy.

We mention this at the beginning of the section.
\end{solution}
\question Section 4.1: please state whether the syntax for
functions in mpmath is identical, or not, to that in Sympy.
\begin{solution}
The mpmath library is not a SymPy submodule, but a separate library,
which we have indicated in the text. Most function names in mpmath are the same as SymPy, but it is not true in every case. We have
added an example to the text showing an instance where they diverge
(summations).
\end{solution}
\question Unless I have
missed something, a weakness of the paper is that speed is mentioned only
twice, on lines 496 and 666. My assumption is that SymPy is slow compared with
its commercial competitors. Please comment on the speed issue, and not just in
the conclusions.
\begin{solution}
  % TODO
\end{solution}
\question Editors are needed for [20].
\begin{solution}
We have done this.
\end{solution}
\end{questions}

@@

\question It is unclear what ``both of these objects'' refer to in line 532. My guess is vectors and dyadic objects, but this should be reworded to make it more clear.
\begin{solution}

\end{solution}

\question In lines 543 and 545, ``rad'' should be removed. Radians are assumed when no other measure (like degrees) is given.
\begin{solution}
These units were removed.
\end{solution}

\question In lines 567--568, ``performing symbolic quantum mechanics'' makes no sense. This should probably be ``computations'', ``solving problems related to'', etc.
\begin{solution}
This was reworded to:
\begin{quote}
to solve problems in quantum mechanics
\end{quote}
\end{solution}

### 2016-08-20 numerics.tex
More consistent use of "module", "submodule", and "package"
Add mpfr reference, this address editor comment 4
FYI
https://groups.google.com/d/msg/sympy/f4t2ORLEKMM/Yu1Y1VcN5ewJ

The implementation of arbitrary-precision floating-point arithmetic is
supplied by the mpmath library~\cite{mpmath}. Originally, it was developed as a SymPy
module but has subsequently been moved to a standalone pure-Python package.
The basic datatypes in mpmath are \texttt{mpf} and \texttt{mpc}, which
respectively act as multiprecision substitutes for Python's \texttt{float} and
\texttt{complex}. The floating-point precision is controlled by a global
context:

@@

Like SymPy, mpmath is a pure Python library.
Internally, mpmath represents a floating-point number
${(-1)}^s x \cdot 2^y$ by a tuple $(s, x, y, b)$ where
$x$ and $y$ are arbitrary-size Python integers
and the redundant integer $b$ stores the bit length of $x$ for quick access.
If GMPY~\cite{GMPY} is installed, mpmath automatically uses
the \texttt{gmpy.mpz} type for~$x$, and GMPY methods
for rounding-related operations, improving performance.

>>>>

The implementation of arbitrary-precision floating-point arithmetic is
supplied by the mpmath library~\cite{mpmath}. Originally, it was developed as a SymPy
submodule but has subsequently been moved to a standalone pure-Python package.
The basic datatypes in mpmath are \texttt{mpf} and \texttt{mpc}, which
respectively act as multiprecision substitutes for Python's \texttt{float} and
\texttt{complex}. The floating-point precision is controlled by a global
context:

@@

Like SymPy, mpmath is a pure Python library.  This is it's major advantage
over the GNU MPFR~\cite{Fousse:2007:MMB:1236463.1236468},
which is much more feature-rich and fast.  Internally, mpmath represents
a floating-point number ${(-1)}^s x \cdot 2^y$ by a tuple $(s, x, y, b)$ where
$x$ and $y$ are arbitrary-size Python integers
and the redundant integer $b$ stores the bit length of $x$ for quick access.
If GMPY~\cite{GMPY} is installed, mpmath automatically uses
the \texttt{gmpy.mpz} type for~$x$, and GMPY methods
for rounding-related operations, improving performance.
